
% introductory paragraph
\lettrine[lines=2,nindent=0pt]{\color{darkgray}\textbf{F}}{ootball} has grown from being the favorite sport in large parts of the world into a respected industry as reflected by high attendance rates and considerable revenues.\footnote{Except where otherwise noted, football refers to the sport also known as soccer.} For instance, the `Bundesliga', the German top league, is a major reason for public mass gatherings and had with 42,700 spectators the highest average attendance per match across all European leagues in the season 2017/18. Also economically speaking, football is synonymous with big business as the `Bundesliga' ranges in second place in Europe with generated revenues of 3.17 billion euros \citep{deloitte2019report}. With such a great appeal, the matches of the football industry generate external effects in many dimensions. On the positive side, there are higher expenditures for consumption (e.g. merchandising, catering, and accommodation) and ultimately increased local tax revenues. On the negative side, the games induce air and noise pollution, together with a higher frequency of accidents, among others. Violent fan behavior is a particularly pervasive aspect in this context and affects individuals' health, the penal system, and the police. In the season 2017/18, the State and Federal Police accumulated over 2.11 million working hours to guarantee the spectators' safety surrounding German professional football matches \citep{zis17_18}. There has been a widely noticed legal dispute between the Federal State of Bremen and the German Football League (DFL) about the reimbursement of costs for high-risk games, in which the \cite{fac_2019} declared the charges imposed by Bremen as legitimate. This study does not only focus on football matches, where social interactions lead to violence, but also deals with the question that is at the heart of the discussion of whether the negative externalities of football games that bring so much prosperity should be absorbed. 




% STRATEGY & DATA
The aim of this paper is to undertake an in-depth examination of the effects that football games in the framework of the German professional football league system exert on violent behavior. To get at the causal impact of football matches on physical assaults, I employ a generalized difference-in-differences approach that exploits the variation in the timing of the football games. I compare the level of assaults on days with and without home games while controlling for any potential source of heterogeneity across days of the week, month, and year, while also accounting for other possible confounding variation from weather and holidays. In order to use the time-series and cross-sectional variation in the data, I construct a panel at the municipality-day level. I collect and match data from various sources to analyze the effect of 4,461 football games over the period from 2011-2015. First, detailed information on all matches in the first three leagues is obtained via web scraping.\footnote{This includes, but is by far not limited to the time and location of games.} Second, the primary outcome variable, the rate of physical assaults, comes from powerful registry data provided by the Federal Criminal Police Office. It covers all individuals that fall victim to a crime against their legally protected rights in the sample regions. Third, data from weather monitors and a time-series of holidays on the state level are matched to the data to account for possible confounders. Last, population figures from the Federal Statistical Office are merged to construct regression weights and build the outcome variable. 



% findings
I find significant and robust evidence that football matches lead to an increase in the rate of physical assaults. A home game increases the assault rate by, on average, 21.5 percent. The effects are almost entirely driven by male victims and are most substantial for the age group 18-29 years. Besides, the impact of football matches is stemming from victims with no prior relationship to the suspect and completed offenses (i.e. no attempts). Assaults on police officers on duty account for 16 percent of the baseline effect found on the entire sample. I find no evidence that the effects of football games on assaults are offset by reductions on days adjacent to the matchday or in nearby areas. The focus on home games in the empirical framework does not pose a threat to the identification strategy, as away games do not change the level of physical assaults in the home district.



% channels 
There are a couple of theories rationalizing spectator violence. For instance, the frustration-aggression hypothesis, proposed by \cite{dollard1939frustration}, interprets violent behavior as an act to rehabilitate self-esteem after an instigating frustrating event, such as the defeat of one's favorite team. Another prominent theory is the idea of social learning, in which spectators mimic the behavior displayed on the pitch \citep{bandura1973aggression, bandura2007social}. In my analysis, I test whether the frustration-aggression hypothesis can serve as an explanation for why there are more physical assaults. To do so, I use emotional cues from either emotionally unsettling events during the game or when game outcomes do not align with pregame expectations. In both cases, there is no evidence that would support the frustration-aggression hypothesis. In the second step, I consider game and team prominence as potential channels. I find a three times higher impact on the rate of physical assaults during high-rivalry matches. This higher effect can be associated with mutual dislike of opposing fan groups or by the social learning theory, as high-rivalry matches exhibit increased game intensity on the pitch. Furthermore, I find larger effects for the highest tier of the league system, which attracts the most people. Ultimately, I cannot pinpoint the one channel at play, but there is suggestive evidence that spectators mimic players' behavior, they select into specific matches, and eventually a mechanic agglomeration effect. 



% back of envelope calculations
Back-of-the-envelope calculations based on my estimates indicate that the crime externalities of football games are substantial. For example, football games in the first three leagues of the German Football league system explain 17.7 percent of assault reports and lead to 18,770 additional assaults in the season of 2014/15. \cite{glaubitz2016kostet} estimate the social cost of an assault at 5,067 euros (in 2020 prices), which implies an annual social cost of 95 million euros.\footnote{The estimated social costs by \cite{glaubitz2016kostet} are conservative compared to other estimates from around the globe. In the United States, the social costs of an assault range from 17,300 to 68,000 euros \citep{miller1996victim,cohen2004willingness}. In New Zealand the social costs are estimated at 6,400 euros \citep{roper2006estimating}, and in Great Britain the estimated costs are 2,300 euros \citep{dubourg2005economic}. All costs are in 2014 prices.}



% Literature
This paper complements the literature on sporting events and their effects on different dimensions of crime. Studies in the US American context often use offense reports from the National Incident Based Reporting System to investigate the impact of American (college) football games on crime. \cite{rees2009college} exploit within agency variation to study the effects of Division I-A college American football games on various offense categories for the years 2000-2005.\footnote{Almost all of the following studies exploit within law enforcement agency variation over time while controlling for weather, holidays and other sources of heterogeneity across time.} They find that assaults, vandalism, arrests for disorderly conduct, and alcohol-related offenses increase in hosting municipalities on game days. More sizable effects are associated with upset game outcomes, defined as when lower-ranked teams win over better-ranked teams. The effects of away games are not significantly different from zero. \cite{lindo2018college} examine the effect of college party culture surrounding Division 1 American football games on sexual assaults. They show that the daily reports of rape victimization among 17-24-year-old women increase by 28 percent on games days. The effects are larger for home games, prominent matches, and teams playing in the better ranked subdivision of Division 1. Furthermore, they show that game outcomes matter: upset wins increase rape victimization strongly. \cite{card2011family} analyze the impact of emotionally unsettling events associated with wins and losses of professional American football teams on family violence for the years 1995-2006. They find an increase of intimate partner violence by ten percent in the event of upset losses (when the home team was expected to win). There are no effects for upset wins or when the game expectations are close. The effects of male-on-female intimate partner violence are pronounced around the end of the game and are larger for more prominent games. In the European context, researchers focus on football matches to estimate the effect of large scale sporting events on crime. \cite{marie2016police} investigates the effect of football matches on crime in London using hourly offense data obtained from the Metropolitan Crime Statistics System. His results show that property crimes increase (fall) by 4 percent (3 percent) for every extra 10,000 spectators attending a home (away) game. Violent crimes are only affected by derby matches. \cite{montolio2016time} study the temporal impact of football matches on criminal behavior in Barcelona (2007-2011). They match reports on registered crime with football matches played by the Football Club Barcelona (FCB) to see whether the games displace criminal activities temporally. They exploit a panel approach in which they compare crime for the same time window on the same weekday, with and without FCB games. Their results indicate temporal displacements for criminal activities of thefts, criminal damage, robberies, and gender violence. Moreover, instances of gender violence increase after home defeats. In a follow-up study, \cite{montolio2019measuring} investigate the spatial dimensions of the crime externalities stemming from football games. Their findings show increased levels of thefts (mainly pickpocketing) across the entire city in the event of a home game. The impact is larger for regions in close proximity to the stadium. During away games, by contrast, the number of thefts around the stadium is relatively smaller. The effects of football matches on assaults are analogous to thefts. In fact, they find a displacement effect of violent fans to areas close to the stadium on days with home matches.
%\footnote{In addition to the papers presented in this section, other noteworthy studies in that strand of research are: \cite{kalist2016national}.}




% extension literature
The empirical analysis in this paper expands previous literature on many dimensions. I assemble a unique data set from various sources to characterize in full the impact of football matches on criminal behavior in the top three German professional football divisions. Using the universe of criminal offenses, I am able to show that almost one in five physical assaults can be attributed to professional football games. In addition to the overall effect, I show effect heterogeneity by victim and crime characteristics. When investigating channels, I find no support for the frustration-aggression hypothesis, but evidence for the social learning theory and a selection story in which violent fans self-select into specific matches.





% roadmap
The remainder of the paper is structured as follows. The next section provides information about football games and how they are associated with violent spectator behavior. Section \ref{sec_soc_ext:data} explains the data and variables. Section \ref{sec_soc_ext:empirical_strategy} contains a description of the empirical framework. Section \ref{sec_soc_ext:results} reports results, validity checks, a discussion on potential channels, and robustness tests. Section \ref{sec_soc_ext:conclusion} concludes.






%--------------------------------------------------------------------
% BACKGROUND
%--------------------------------------------------------------------
\bigskip
\section{Background}\label{sec_soc_ext:background}



% German Football League System
\subsection{German Football League System}
The three full professional divisions of the German football league system are held under the jurisdiction of the German Football Association (DFB) and the German Football League (DFL). While the two top flights `Bundesliga' and `2. Bundesliga' are organized by the DFL, the third division `3. Liga' is run by the DFB itself. Teams can move through the levels of the league system by either promotion or relegation. The two top divisions consist of 18 teams that play 17 home and away games in one season. The third level contains 20 teams with 19 home and away games. The empirical approach of this paper exploits the variation in the timing of matches. Since 2006, the preparation of the match schedules for the `Bundesliga' and the `2. Bundesliga' is carried out with a software that uses integer linear programming.\footnote{For details of how the match plans are made, please refer to \href{https://www.bundesliga.com/de/bundesliga/news/so-entsteht-der-spielplan-bundesliga-2-bundesliga-dfl-computer-software.jsp}{https://www.bundesliga.com/de/bundesliga/news/so-entsteht-der-spielplan-bundesliga-2-bundesliga-dfl-computer-software.jsp}} The software outlines the rough details such as the fixtures per matchday. The exact date and time, however, are determined in the course of the season. The late exact scheduling allows to take into account guidelines from local authorities, security bodies, the Central Sports Intelligence Unit (ZIS), international football associations (FIFA/UEFA), fans, clubs, and stadium operators. In addition to obvious restrictions such as the fact that neighboring clubs should play their home matches at a different time, the DFL has to consider public holidays, parallel major events, or match dates of international competitions.



% Soccer and Crime
\subsection{Football and Violent Crime}

% ULTRAS & HOOLIGANS
Spectator violence has a long tradition in the environment of professional football in Germany. From the change of terms for football fans you can see that spectator behavior underwent quite a transformation \citep{pilz2005kuttenfan}. In the 1960s and 1970s, the peaceful fan base was specified as `camp-followers', while one decade later the first problems of spectator violence arose with the presence of `football rowdies'. In the 1980s, spectator violence was particularly omnipresent with the hooligans' movement. Since the end of the 1990s, a new group appeared in the stadiums, the `Ultras'. Originated in Italy, the `Ultras' have committed themselves to fight against the commercialization of football and to revitalize traditional football culture. Over the last years, the number of violent fans has been increasing. The police differentiate between three types of football fans. Category A consists of peaceful fans, category B includes violence-prone fans, and category C contains fans that actively seek violence (violent criminals). Originally, the `Ultras' could be located mostly in category A and sporadically in category B. Recently, however, a substantial share of the `Ultras' has been classified as members of categories B and C. For instance, the \cite{zis17_18} identified 13.633 individuals belonging to either violence-prone or violence-seeking fans in the top three leagues for the season 2017/18. With more potential agitators, police officers report more aggressiveness directed against themselves \citep{feltes2010fussballgewalt}. Yet, it is not only the problem of having more violent fans, but also that the event character of violence during football games is on the rise \citep{pilz2005kuttenfan}. Violent actions are increasingly less connected with the events on the pitch. In fact, violence serves as an opportunity for identification in the quest for one's identity. Some violent fans find self-assertion by joining forces with like-minded individuals and discover their strength. The question is how to address the challenges coming from `hooligans' and `Ultras'. \cite{poutvaara2009police} investigate the impact of counter-hooliganism policy in the Swedish context. They find that the reallocation of the Sport Intelligence and Tactical Unit to other activities was accompanied by increases in violent crime by hooligans. \cite{feltes2010fussballgewalt} emphasizes the concept of `balanced policing', which acknowledges that risks coming from the spectators are dynamic and require quick responses to either escalate or deescalate.\footnote{Depending on which types of violent fans the police is facing, the officers have to adjust their behavior accordingly \citep{feltes2010fussballgewalt}. On the one hand, hooligans require a zero-tolerance policy and any absence of drastic measures is seen as an invitation to live out violent desires. On the other hand, Ultras should be given some leeway to allow for self-regulation.} 



% THEORIES OF SPECTATOR VIOLENCE
There are a couple of theories rationalizing spectator violence.\footnote{Theoretically, there could also be a decreasing assault rate due to football games. This may happen if individuals with a higher propensity to act violently attend the matches. \cite{dahl2009movie} find a similar effect for violent movies and name this a `self-incapacitation effect'. Yet, as only hourly data can reveal the effect and the analysis draws on daily data, I focus on theories of why the rate of violence is increasing due to football games.} First, the higher number of assaults could be a mere mechanic effect arising from increased social contacts and a change in how the interactions take place \citep{montolio2019measuring}. For instance, the `Bundesliga' has the highest average attendance per match across all European main football leagues \citep{wicker2017effect}. In the season of 2014/15, more than 42,000 spectators attended on average a game in the highest flight of the German league system. This aspect is particularly relevant for smaller municipalities, which attract thousands of spectators per game \citep{lindo2018college}. Second, spectator violence may be the result of physiological arousal \citep{branscombe1992role}. Potential determinants of physiological arousal are increased heart-rate, blood pressure, and respiration, which may influence affect, cognition, and (anti-)social behavior. Third, \cite{dollard1939frustration} propose the idea of the frustration-aggression hypothesis.\footnote{\cite{berkowitz1989frustration} presents a good overview of the frustration-aggression hypothesis.} If the favorite team suffered a defeat, this may precipitate aggressive fan behavior (e.g. see the results obtained by \cite{card2011family}). Fourth, \cite{bandura1973aggression, bandura2007social} postulates the notion of `social learning' or mimicking behavior. Independent of the outcome of the match, the sheer observation of the game is enough to unleash violent actions.
%Second, instinctual theories of aggression find human aggression genetically based \citep{baron1977human}. In other words, aggressive behavior is an inevitable predisposition. 



% THE ROLE OF ALCOHOL
Alcohol plays a crucial role in the context of spectator violence in German football stadiums. \cite{cook2013virtuous} describes the pharmacological effects of drinking alcohol on aggression and cognitive functioning. With drinking comes a loss of inhibition and impairments of judgment. Furthermore, participants of experiments have been found to show more aggressive behavior after drinking. The 2018 edition of the \citeauthor{PCS2018} specifies that more than every fourth assault (26.2 percent) was committed under the influence of alcohol. While some countries banned alcoholic drinks within the premises (e.g. Brazil since 2003), the rules in German stadiums are somewhat ambiguous. The DFB's security guidelines stipulate that the sale of alcoholic beverages is forbidden before and during games in the stadium. Nevertheless, with the approval of the responsible local security organs, the hosting clubs may, on their own authority, deviate from the regulations. Only in the case of high-risk games, the clubs are urged to adhere to the ban.\footnote{For details, please refer to: \href{https://www.sueddeutsche.de/sport/fussball-alkohol-in-stadien-nur-mit-ausdruecklicher-einwilligung-dpa.urn-newsml-dpa-com-20090101-150625-99-01743}{https://www.sueddeutsche.de/sport/fussball-alkohol-in-stadien-nur-mit-ausdruecklicher-einwilligung-dpa.urn-newsml-dpa-com-20090101-150625-99-01743}.} The clubs, however, have a strong incentive to deviate from the ban as more than one-sixth (538 million Euros in the season of 2017/18) of the `Bundesliga' club's revenues comes through earnings on the matchday (e.g. tickets and catering) and the sale of alcoholic drinks contributes to that amount substantially \citep{deloitte2019report}. For this reason, alcohol and the potential side effects are very present in German football arenas.
% any data on how much is consumed? https://www.news.de/sport/855220143/schalke-hat-die-durstigsten-fans/1/
% Statista survey: 56  percent of respondents of a panel survey were in favor of banning alcohol from football stadiums, 35 percent defended the right to drink alcohol in football stadiums: https://de.statista.com/statistik/daten/studie/548802/umfrage/verbot-von-alkohol-in-fussballstadien/





%--------------------------------------------------------------------
% DATA & VARIABLES
%--------------------------------------------------------------------
\bigskip
\section{Data}\label{sec_soc_ext:data} 
% Description data
The data set used for the analysis covers the time window 2011-2015 and contains regions in which professional football games take place. The analysis is carried out on the municipality level, the lowest territorial division in Germany. I combine various data sources in order to examine the impact of professional football games on violent behavior.



\subsection{Crime data}
The crime data comes from the German Police Crime Statistics, which is provided by the Federal Criminal Police Office.\footnote{Many aspects of the data preparation are inspired by \cite{hener2019noise} who is using the same data to examine the causal effect of noise pollution on criminal activities.} It covers the universe of individuals that fall victim to a crime against their legally protected personal rights between 2011 and 2015. Yet, as the data is only reported after the end of police procedures, the data is truncated to the period from January 2011 to May 2015 to avoid any problems with lags between the occurrence of the crime and when it is reported. Besides, the months of June are excluded from the analysis as no games are played during that time of the year. The data includes next to the time and place (municipality level) of the crime, the crime type code, the victim's age and gender, characteristics of the assault execution (attempt/completed act, usage of a firearm, lone operator/crime was committed by a group) and information on the relationship between victim and suspect.\footnote{The relationship between victim and suspect is retrieved in two ways. On the one hand, this covers formal relationships (such as types of kin or acquaintances). On the other hand, this examines relationships in a spatial-social manner (for instance living in the same household, or being in an education or care context).} Roughly 40 percent of the victims are female, the average age is 32 years, and 40 percent of the victims haven't had any prior relationship with the suspect.
% (up to the hour) 


% crime - assault distribution over time
\input{soc_ext_fig__assault_time_distribution_2014}

The micro-data is aggregated to the municipality-day level and the main outcome, the assault rate, is defined as the number of assaults per million population. Assaults are defined as actions involving physical violence. For that purpose, I use the crime type code of 'intentional simple bodily harm' (\textit{Vorsätzliche einfache Körperverletzung} § 223 StGB). There are roughly 120 types of criminal offenses in use (recorded in 6 digit codes), though the vast majority of the cases concentrate on only a handful of codes.\footnote{The top 10 of the most prevalent crime keys comprise more than 90 percent of the cases.} Figure \ref{fig_soc_ext:offense_types_distribution_2014} in the Appendix shows the distribution of cases per crime key for the twenty most common offense types in 2014. The intentional simple bodily injury is by far the most prevalent offense and it contributes to about 45 percent of all cases. The prevalence of the offense type and its association with aggression motivates the decision to have assaults in this definition as the main outcome variable.

Figure \ref{fig_soc_ext:assault_time_distribution_2014} shows the distribution of assaults over time. Panel A displays the variation of assaults per hour of the day.\footnote{Roughly 15 percent of the observations do not contain information on the hour. This does not entail any consequences for the main analysis, as I examine daily variation in the assault rate.} The number of assaults is increasing over the course of a day and peaks around midnight. In order to assign the cases that occur during the early hours to the day, in which they are rooted in, I define a day as running from 6:00 AM until 5:59 AM. Panel B shows the distribution of assaults across days of the week. There are relatively more assaults on Friday and Saturday, whereas the other days have assault rates that are somewhat smaller. Panel C shows that the number of assaults has a strong seasonal pattern with the highest value recorded in May and the smallest figure in August.\footnote{Panel C shows the number of assaults across months of the year while adjusting for the number of days per month.} Panel D confirms this impression by plotting the daily number of assaults. New year's eve is a particularly impressive outlier in this context.\footnote{Panels C and D are showing data for the year 2014 only. Appendix Figure \ref{fig_soc_ext:assault_year_distribution} shows the distribution per day for different years.}

% XXX per day roughly around 320 per day, peak on new year's eve, high on holidays (carneval) 

% MAP : (I) WEATHER monitors & (II) NEIGHBOUR regions
\input{soc_ext_fig__map_gem_stadiums_monitor_and_neighboring_regions}



\subsection{Football data}
The data on football matches is self-collected and is obtained via web scraping from kicker.de and transfermarkt.de. It captures all matches that are carried out in the context of the first three leagues of the German football league system over the period January 2011 until May 2015. The data contains detailed match and table standing parameters, e.g. time and place of the match, attendance, pregame point difference, goals, penalties, cards, referee characteristics, among others. Furthermore, there is comprehensive information on the single teams; such as the size of the team, average age, market value, and the number of foreign players. The stadiums where the matches take place are geographically encoded and will \textit{activate} the treatment in the region in which the stadium is located. Figure \ref{fig_soc_ext:map_gem_stadiums_monitor_and_neighboring_regions} depicts a map with all 69 stadiums in the data set. 

% XXX stadiums in which not at least a full season is played are excluded from the analysis.\footnote{This affects the stadiums in Lübeck (1), Reutlingen (12), Elversberg (14), and the Air-Berlin stadium in Dusseldorf (3). The number in parenthesis indicates the number of games that were held in that venue. This excludes 0.6 percent of all matches.}

% Descriptives about football matches
\input{soc_ext_fig__descriptives_matches_time_attendance}

Figure \ref{fig_soc_ext:descriptives_matches_time_attendance} illustrates some insights into key variables. Panel A shows the number of matches per day of the week and league. The vast majority of the matches is taking place between Friday and Sunday. This is particularly true for the highest league, whereas lower leagues play occasionally on weekdays, too. Matches happening on weekdays are exclusively held in the evenings. In contrast, matches on weekends occur usually in the afternoon. The inclusion of day-of-week fixed effects in the baseline specification helps to account for the higher share of games played on weekends, which are associated with higher levels of criminal behavior. The number of spectators vary substantially across the three professional leagues, as depicted in Panel C. The `Bundesliga' draws the highest attention with an average audience of 44,000 people per game, followed by the second league with an average 17,000 fans per match, and the lowest league attracts on average a bit less than 6,000 fans per game.

When investigating channels of how football games may affect assaults, I exploit betting odds that were obtained via web scraping from oddsportal.com. The betting odds give an idea about pregame expectations. I translate the odds of the three game outcomes to probabilities, which are the inverse of the betting odds. The probabilities serve as good predictors for game outcomes, as it is shown in Appendix Figure \ref{fig_soc_ext:pregame_probability_spread_outcomes}.\footnote{Panel A of Figure \ref{fig_soc_ext:pregame_probability_spread_outcomes} shows the close relationship between the realized score differential and the probability spread. Panel B and C demonstrate that the probability of victory is increasing with higher values of the probability spread.}


\subsection{Weather data}
The weather data comes from Germany's National Meteorological Service (\textit{Deutscher Wetterdienst}). In order to construct the weather control variables, I use the sample of weather monitors, which measure relevant weather variables over the sample period.\footnote{I use the following weather variables as daily averages: the daily average, minimum and maximum air temperature, minimum ground temperature, vapor pressure, air pressure, cloud cover, air humidity, precipitation, sunshine duration, snow depth, and wind velocity.} From this set of monitors, I choose the weather monitor with the closest proximity to a stadium. The assigned monitor-stadium pairs can be found in Figure \ref{fig_soc_ext:map_gem_stadiums_monitor_and_neighboring_regions}. The quality of the matches between weather monitors and stadiums is high, as the average distance between stadiums and monitors is 15 kilometers.\footnote{The distribution of the distance of monitor-stadium pairs can be found in Appendix Figure \ref{fig_soc_ext:hist_distance_stadium_monitors}.} Few of the weather variables have missing data, which are filled by propagating the last valid observation forward to the next valid observation (i.e. forward fill).\footnote{The weather variables with missing data are (with the share of missing data in parenthesis): Cloud cover (<1.2 percent) and snow depth (<0.6 percent).}



\subsection{Holidays}
In order to capture any variation of the crime rate between ordinary and special days, I add controls for public and school holidays, which may differ on the state level.\footnote{The data for the school holidays comes from `The Standing Conference of the Ministers of Education and Cultural Affairs of the Länder in the Federal Republic of Germany' (\textit{Kultusminister Konferenz}). Data on public holidays are collected from \href{https://www.schulferien.org/deutschland/feiertage/}{https://www.schulferien.org/deutschland/feiertage/}.} Furthermore, I add a dummy variable for peculiar days (New Year's Eve and the days surrounding Carnival), which are not holidays, but certainly shift the crime rate anyways.



\subsection{Regional Database}
The Federal Statistical Office and the statistical offices of the Länder provide a database of detailed statistics by various subject areas on a very granular spatial level. Thus, I am able to build a panel on the municipality-year level containing comprehensive information on topics such as territory, population, labor market, election results, housing, economic sectors, and public budgets. The information is used to construct weights for the regression analysis or to build assault rates. {\color{red}Furthermore, the plan for the future is to leverage the high dimensionality of the data in order to investigate treatment effect heterogeneity with the help of machine learning algorithms (e.g. by making use of causal forests).} 







%--------------------------------------------------------------------
% IDENTIFICATION
%--------------------------------------------------------------------
\bigskip
\section{Empirical strategy}\label{sec_soc_ext:empirical_strategy}

In order to identify the causal effect of football matches on criminal behavior, I exploit within region variation over time. To be precise, I compare the regional assault rate on the game day to the expected assault level conditional on the day of the week, month, and year while additionally accounting for other possible confounding variation stemming from weather and holidays. In other words, the counterfactual regional assault rate (what would be expected on a game day in absence of the game), e.g. a Saturday in April 2012, is obtained by using the regional assault rate on other Saturdays in April 2012, where no games take place. %In other words, the regional assault rate on a given date without a game (say a Saturday in April 2012) serves as counterfactual for what would be expected on a game day (on a Saturday in April 2012) in absence of the game. 


The identification strategy is based on a generalized difference-in-differences approach to study the impact of a football match on violent behavior. Let Assaults$_{rdmy}$ denote the assault rate in region $r$, on day-of-the-week $d$, in month $m$ and year $y$. The assault rate is defined as the number of assaults per million population and is given by:

\begin{align}
\text{Assaults}_{rdmy} = \alpha + \beta\ (\text{Gameday}_{rdmy}) + \vartheta_r + \underbrace{\gamma_d + \eta_m + \theta_y}_{\text{date}_{dmy}} + \lambda \mathbf{X}_{rdmy} + \varepsilon_{rdmy}
\label{eq_soc_ext:model}.
\end{align}

$\text{Gameday}_{rdmy}$ is a binary variable equal to unity when a home game takes place and zero otherwise. Region fixed effects $\vartheta_r$ capture time-invariant differences between regions and ensure that the identification is driven by within instead of between region variation over time. The vector $\text{date}_{dmy}$ contains fixed effects for day-of-the-week ($\gamma_d$), month ($\eta_m$), and year ($\theta_y$). This way, the model flexibly controls for day-of-week specific heterogeneity, seasonal effects, and long-run time trends. I expand on the baseline model by adding interactions of region fixed effects with all elements of $\text{date}_{dmy}$, i.e. region-by-day-of-week fixed effects, region-by-month fixed effects, and region-by-year fixed effects. The interactions account for systemic changes in the degree of violent behavior over the year for each region. The vector $\mathbf{X}_{rdmy}$ includes school and public holidays and weather controls.\footnote{Public holiday controls include binary variables (on the level of the Federal States) for All Saints, Ascension Day, Assumption Day, Christmas, Corpus Christi, Epiphany, Easter, German Unity Day, Good Friday, Labor Day, New Year, Penance Day, Pentecost, and Reformation Day. Moreover, it contains dummy variables for Carnival and New Year's Eve.\newline Weather controls (on the regional level) include average air temperature, maximum air temperature, minimum air temperature, minimum ground temperature, steam pressure, cloud cover, air pressure, humidity, average precipitation, sunshine duration, snow depth, and wind speed.} I compute two-way cluster-robust standard errors to capture arbitrary correlation at the region-year and year-month levels. Observations are weighted using population figures from the Federal Statistical Office.


% displacement effects - temporal & spatial
The implicit assumption to interpret the parameter of interest $\beta$ as the causal effect of a home game on violent behavior is that the location and the time of a football match are orthogonal to the number of assaults, conditional on the covariates. However, displacement effects may pose a threat to identification. On the one hand, this refers to spatial displacement effects, which may occur when (violence-prone) populations come from distant regions to the game. On the other hand, this includes temporal displacement effects, which happen when assaults from neighboring days are shifted to the game day. In both cases, the parameter would overestimate the impact of a football match on violent behavior as the offense would have been committed anyway, but just at another time or in another place. To rule out the possibility that displacement effects pose a threat to the validity of the identification strategy, I investigate the effect of football games on neighboring regions and days surrounding game days in section \ref{sec_soc_ext: threats}. I find that the main results, an increase in violent behavior in regions where football games take place, are not neutralized by decreases in adjacent regions or on days surrounding game days.


% control group problem - away games
Given the setup of the empirical approach, there could be another threat to the validity of the identification strategy. By focusing on home games in the main analysis, the counterfactuals could potentially be biased downwards as days with away games are part of the control group. This control group problem might stem from the fact that violent fan groups travel with their team on away games and the assault rate in the home region might consequently decline. In order to alleviate this concern, I redo the analysis and differentiate between home and away games. When allowing for distinct effects for home and away games, I find that away games do not significantly affect the assault rate in the home region. 









%--------------------------------------------------------------------
% RESULTS
%--------------------------------------------------------------------
\bigskip
\section{Results}\label{sec_soc_ext:results}



% MAIN RESULTS
%--------------------------------------------------------------------
\subsection{Main results}

% fig: intuitive representation: assrate across DOW
\input{soc_ext_fig__assault_rate_across_dows}

% descriptive figure, ass_rate across DOW
Before presenting the regression results, Figure \ref{fig_soc_ext:assault_rate_across_dows} gives an intuitive preview of the main findings. Using the same data as the main analysis, it shows the assault rate across days of the week.\footnote{Appendix Figure \ref{fig_soc_ext:assaults_across_dows} shows the same scheme when using the raw number of assaults across days of the week for days with and without games.} The daily rates are presented for days with and without home games. The average daily assault rate is higher for weeks in which there is a home game compared to when no game takes place. The difference in means is statistically significant for all days except for Tuesdays and Wednesdays.\footnote{The average assault rate on Thursdays with games is very high in relation to Thursdays without games. This may be the result of very few games taking place on a Thursday. In the dataset, only 18 (0.4 percent) games take place on a Thursday.} The empirical model exploits the variation in assaults across days of the week, and particularly how the pattern varies between weeks with and without games.

% main result table with more controls
Table \ref{tab_soc_ext:reg_fe_assrate} reports estimates corresponding to equation \ref{eq_soc_ext:model} when adding progressively more controls. The dependent variable is the assault rate, which is defined as the number of offenses per million population. In column 1, I include region, day-of-week, month, and year fixed effects. Column 2 adds weather controls. Holidays fixed effects are included in column 3. Column 4, controls for region-specific date fixed effects. Note that although the estimates vary marginally across columns, I use throughout the paper the model presented in column 4 for subsequent analyses. The estimate from the preferred specification in column 4 suggests that a home game increases the assault rate by an average 2.677 assaults per million population. This corresponds to an increase of 21.5 percent from the game-free level.

% main regression table: include fixed effects
\input{soc_ext_tab__reg_fe_assrate}


% table effect heterogeneity: victim and crime chars
\input{soc_ext_tab__reg_fe_asrate_victim_crime_chars}

% victim characteristics
After establishing that football games lead to more assaults, I shed light on who are the victims of these additional offenses. Panel A of Table \ref{tab_soc_ext:reg_fe_assrate_victim_crime_chars} shows the effect of football games on the assault rate by victim characteristics. First, Columns 1 and 2 show the effect heterogeneity by gender. Although the estimates for women and men are statistically different from zero, the large majority of the additional victims are male. Male victimization rates increase on average by 2.432 offenses per million population. The increased assault rate for males consumes more than 90 percent of the effect found on the entire sample. Figure \ref{fig_soc_ext:fe_age_profile} shows for each gender the age profile of the impact of football games on the assault rate. For women, the point estimates are small in magnitude and not significantly different from zero. The one exception is women aged 40-49 years. In contrast, the effects of adult men are throughout significant. The largest effect for males is found in the age bracket 18-29 years and is decreasing thereafter. Second, I explore effect heterogeneity by the relationship between victim and suspect. The columns 3 and 4 distinguish the relations from a formal perspective, such as kin or acquaintances. Although both estimates are positive and statistically significant, the majority of additional assaults is coming from victims who have not had any prior connection to the suspect. The victimization rate of strangers to the suspect rises by on average 1.939 assaults per million population. This implies that almost three out of four extra cases originate from this type of victim-suspect pair. Column 5 considers spatial-social relationships, in this instance whether victim and suspect live in the same household. A football game increases domestic assaults by on average 0.096 assaults per million population. This implies that only few (13 percent) victims that have known the suspect also live in the same household. Third, the effect of the victim's occupation is investigated. Column 6 shows the impact of football games on violent behavior experienced by police officers on duty. The victimization rate of police officers increases on average by 0.434 assaults per million population. Thus, almost 16 percent of the additional assaults due to football games can be attributed to attacks on police officers. Taken together, the additional victims due to football games are young males with no prior relationship to the suspect. This description of the victims has demographic similarities with the population who usually gathers in and around football stadiums \citep{pwc2016}.

% age profile of effect, ba gender
\input{soc_ext_fig__fe_age_profile}

% crime characteristics
In the next step, I consider effect heterogeneity by crime characteristics. Columns 1-4 in Panel B of Table \ref{tab_soc_ext:reg_fe_assrate_victim_crime_chars} show the heterogeneity of the effect by the time of the offense. The impact of a football game on the assault rate is smallest in summer months with 1.482 assaults per million population, corresponding to an increase of 13.4 percent. The summer months (July and August) are the time when the football season starts and results are not yet so important. The estimates are higher (but not significantly different from each other) during other times of the season. Columns 5 and 6 consider effect heterogeneity by assault execution (attempt/completed offense). I find that the effects of football games on the assault rate are mostly driven by completed offenses with on average 2.310 additional assaults per million population. The share of attempted assaults is increasing, too, but at a smaller rate with on average 0.367 additional attempted assaults per million population.




% Threats to validity
%--------------------------------------------------------------------
\subsection{Potential threats to Identification and Validity of the Design}\label{sec_soc_ext: threats}
This section considers the possibility that I measure not additional assaults due to football games but merely displacements of offenses. Furthermore, the sensitivity of the main results is tested by additionally including away games into the model.

% Intro why we see displacement
Previously, I showed evidence that football games increase the assault rate. However, it could well be that I only pick up an effect that displaces offenses. For example, the increase in violent behavior may be offset by a reduction in assaults in other areas or at different times \citep{lindo2018college}. In other words, the assault would have been committed anyway, but just at a different time or another location. One potential explanation for that are different population flows around days on which games take place.

% table  SPATIAL & TEMPORAL DISPLACEMENT 
\input{soc_ext_tab__reg_fe_assrate_displacement}

% spatial displacement
In order to estimate \textit{spatial displacement effects}, I investigate the impact of football matches on nearby regions. I define neighboring regions to constitute nearby regions. A neighboring region is a municipality that shares a common boundary with a region that contains a stadium.\footnote{If two municipalities share a common border and each of the regions contains a stadium, both regions will not serve as neighbor regions and they are dismissed from the set of spatial spillover candidates.} Figure \ref{fig_soc_ext:map_gem_stadiums_monitor_and_neighboring_regions} shows a map of the selected regions. The sample of neighboring regions exhibits a considerably higher number of observations. This is owed to the fact that a region with a stadium has, on average, a bit more than 11 neighbor municipalities. Panel A of Table \ref{tab_soc_ext:reg_fe_assrate_displacement} shows the estimates of the impact of a home game on these neighboring regions. In comparison to the baseline effects, the spatial spillover coefficients are small and not significantly different from zero. Consequently, the results do not suggest that offsetting spatial spillover effects are present.

% temporal displacement
In the next step, \textit{temporal displacement effects} are being considered. To capture this possibility, I include one-day leads and lags of the game day indicator. Panel B of Table \ref{tab_soc_ext:reg_fe_assrate_displacement} contains the estimates when enriching the baseline model with the temporal spillover components. The estimates of the gameday itself are not significantly different from the baseline model. The coefficients of one day before and after the game are almost always small in magnitude and not statistically significant. The only exception is the estimate of the day before the game in column 4, which implies that a football game leads to an average increase of 0.352 assaults per million population on the day before a game. The positive coefficient, however, does not suggest temporal displacement effects, which would require a negative estimate. This rather indicates additional assaults due to activities on the day preceding important games. 

% table  HOME & AWAY games 
\input{soc_ext_tab__reg_fe_assrate_home_away}

% Home and Away games
As discussed above, the baseline model considers the effect of football matches, but for home games only. This restriction could entail problems for the validity of the design. When away matches are not accounted for, they end up in the control group. The control observations might be biased downwards when the local team is playing away and the most devoted (and eventually violent) fans travel with their team out of the municipality. This could mean that the assault rate at home is decreasing due to the absence of local agitators and days with away matches cannot be used as control units anymore. In order to address this concern, I investigate the \textit{effect of home and away matches separately}. To analyze the effect of away matches, I have to build up my data set differently. In the baseline version, the football data is merged with the other datasets on the match level (with the home team's region ID as the identifier). In this case, I use the football data on the table standings level. In other words, both the home and away teams are matched to a region. One consequence of this approach is that it makes the treatment status of single regions less clear.\footnote{When only considering home games, the treatment status is not a problem. This is due to the fact that local authorities do not allow two home games on the same day.} For instance, the treatment status of regions that have more than one team is ambiguous when there is a home and an away match at the same day. In order to alleviate this concern, I drop the third league from the sample and focus only on the first two leagues.\footnote{There remain still a few ambiguities but they are solved in the following way: in 12 percent of the matches there is still a doubling of two teams per region playing on the same day, either one match at home and one game away, or two away games. In the latter case, the status of the region is set to away. In the former case, I put it to home.} This approach helps to clarify the treatment status of a region considerably. Table \ref{tab_soc_ext:reg_fe_assrate_home_away} shows the results when investigating home and away matches separately. To be able to compare the estimated effects, column 1 shows estimates retrieved from the baseline model (home matches only) when adjusting the sample as described above. Column 2 presents the estimates when allowing home and away games to have an impact on the assault rate. The effect of a home game is sizable and leads to an average increase in the assault rate by 2.910 assaults per million population. The coefficient is not significantly different from the one in column 1. A negative and significant away game estimate would pose a threat to the identification strategy. However, the estimate of an away game is small in magnitude and not significantly different from zero. Thus, the results suggest that the focus on home games only does not lead to an invalid identification strategy. 




% Channels
%--------------------------------------------------------------------
\subsection{Channels}
This section investigates potential mechanisms through which football games may lead to additional assaults. I consider emotional cues and game prominence as potential channels.

% table EMOTIONAL CUES
\input{soc_ext_tab__reg_fe_assrate_emotional_cues}

% emotional cues
In the first step, I consider emotional cues as potential channels of how football games can affect assault rates. This is motivated by the findings of \cite{card2011family} who demonstrate that unexpected losses of local football teams spur family violence. The results fit well into the context of the frustration-aggression hypothesis, first proposed by \cite{dollard1939frustration}, which predicts belligerent behavior in the event of frustrating events. \cite{rees2009college} show in a similar fashion that there are more violent offenses when the local football team loses. For this reason, I analyze in Table \ref{tab_soc_ext:reg_fe_assrate_emotional_cues} whether visceral factors may be the reason for the additional assaults. Column 1 investigates whether the mere existence of an emotionally upsetting event during a game leads to a higher assault rate. To answer this question, I build an index that equals one for games in which there occur at least one of the following potentially unsettling events: a penalty is awarded (20 percent of all games), a red card is being issued (10 percent of all games), or the referee receives a non-sufficient grade (15 percent of all games). The index shows that 35 percent of all games experience at least one upsetting episode as defined in the previous categories. The estimates in column 1 do not suggest that emotional cues trigger more violent behavior as the estimates for games with and without upsetting events are not significantly different from each other. 	In the second and third column, I show estimates following the approach of \cite{card2011family}, in which I examine the impact of game outcomes relative to their pregame expectations. Pregame expectations are included in the analysis as matches with contrasting predictions may be very different from each other. By including predicted outcomes, I can estimate the effect when a team is expected to win but loses, and vice versa. Using data from oddsportal.com, I define a game too close to be predictable when the absolute probability difference between win and loss is smaller than 20 percentage points.\footnote{Similar results are obtained when defining other threshold values and when I deviate from the symmetry around the origin.} When the spread's value is exceeding the threshold, the home game is expected to win or lose, respectively. Around 45 percent of the games are expected to be close, another 45 percent to be won, while 10 percent of the games are expected to be lost. The significantly larger share of wins may be attributed to the home-advantage. Column 2 starts out by examining the effects of matches with distinct predicted match outcomes. The estimates do not suggest that the effect of games with different predicted outcomes differ systematically from each other. In column 3, I additionally include interactions between expected and true game outcomes. The estimates are relatively small in magnitude and not significantly different from zero, implying that unexpected wins/losses do not cause additional assaults. Taken all together, there is no evidence that emotional cues drive violent behavior in the context of professional football games in Germany.


% table TEAM/GAME prominence
\input{soc_ext_tab__reg_fe_assrate_prominent_games}


% game/team prominence (rivals & league)
Second, I use information about game and team prominence as potential channels. For many sports fans, the affection for the own football team plays a central role \citep{wann1993sports}. Almost as important as the passion for the own team is the cultivation of animosities against rivals. Often resentments and hatred can translate into violent actions \citep{nassauer2011hate}. In order to test the hypothesis that the additional assaults are the result of hostile feelings, I compare the impact of matches in which known rival teams are playing against each other to regular matches. To answer this question, the game day indicator of equation \ref{eq_soc_ext:model} is replaced by an interaction with a dummy variable that equals one for high-rivalry matches. Local derbies (games between two competing teams that are based in regions of close geographical proximity) constitute high-rivalry matches.\footnote{Appendix Table \ref{tab_soc_ext:prominent_games_derbies} gives an overview of high-rivalry matches.} The estimates are shown in columns 1 and 2 of Table \ref{tab_soc_ext:reg_fe_assrate_prominent_games}. Games that are classified as high-rivalry matches lead to an average increase of 8.320 assaults per million population. This corresponds to more than three times the baseline effect. Although the standard errors are relatively large, as only 2.5 percent of the games are classified as high-rivalry matches, the effect is significantly different from regular matches. The higher rate of physical assaults during high-rivalry matches could also be rationalized by the social learning theory, in which spectators mimic the behavior displayed on the pitch. For instance, the number of cards awarded is significantly higher in high-rivalry matches.\footnote{The number of cards in derby matches is 4.324\ (s.e.=0.174), while for regular matches the number of issued cards is 3.906 (s.e.=0.029).} Next to the explanation of opposing fan groups' mutual dislike or social learning, it could be the mere prominence of teams/games that attract more (violent) people. As shown in Figure \ref{fig_soc_ext:descriptives_matches_time_attendance}, attendance varies substantially across the tiers of the German football league system. Columns 3-5 shows the effect of games by the league in which the teams are playing. The largest effect is in the highest tier, the `Bundesliga', in which a game leads to a rise in the assault rate by, on average, 3.530 assaults per million population. The impact of matches from the first league is significantly larger than the effect of matches from the other two leagues.
Summing up, there is evidence that the prominence of games is a crucial determinant for the effect size of football matches on the assault rate, where high-rivalry matches and games in the top league lead to more assaults.


% Machine Learning application: effect heterogeneity in other dimensions
{\color{red} missing: ML application: do I find evidence for deprived regions?}\newline
sample split: larger effects rural (2x basline), ue rate no effect 










% Robustness
%--------------------------------------------------------------------
\subsection{Robustness tests}
I perform several sensitivity and placebo tests to assess the robustness of the findings. The results of these checks are reported in Table \ref{tab_soc_ext:reg_fe_assrate_robustness}. Overall, the sensitivity tests demonstrate that the main results are robust to alternative specifications and estimations, indicating that football games lead to more assaults.\newline

% table Robustness: ECONOMETRICS & other VIOLENCE
\input{soc_ext_tab__reg_fe_assrate_robustness}


% Alternative econometrics: sample, weights, estimation
\textbf{Alternative econometric specifications.--} First, I show that the results are not sensitive to alterations of the \textit{sample}. Adjustments to the sample may be necessary as some of the games are played on different days than what was originally planned. The deviation from the original match plan may pose a risk to the allocation of games that is plausibly random. For this reason, I drop the set of rescheduled games from the analysis. The results are almost identical to the baseline results.\footnote{From the 4,461 games in the sample, 2.24 percent are rescheduled. The vast majority of rescheduled matches (95 percent) are in league three.} Second, I show the effects when leaving out population \textit{weights} from the regressions. The effect of football matches without population weights is considerably larger: a home game leads to an increase in the average assault rate by 5.342 assaults per million population. The higher coefficient compared to the baseline specification indicates that the effects of rural areas must be larger, as they obtain higher relevance when population weights are omitted. Third, I show that my results are robust to a different \textit{estimation} procedure and an alternative measure of the \textit{dependent variable}. In the following, I use the raw number of assaults as the outcome variable. Because of the discrete nature of the dependent variable, and given that there are many cells with zero reports, a Poisson model is the model of choice. The corresponding regression specification reads as follows:
\begin{align}
E\left[ \text{Assaults}_{rdmy}\right.&|\left.\text{Gameday}_{rdmy},\vartheta_r,\text{date}_{dmy}, \mathbf{X}_{rdmy} \right] \nonumber \\ &= exp\left( \beta\ (\text{Gameday}_{rdmy}) + \vartheta_r + \text{date}_{dmy} + \lambda\mathbf{X}_{rdmy} \right). 
\label{eq_soc_ext:poisson_model}
\end{align}

Due to the nonlinearity of the model, the coefficient in Table \ref{tab_soc_ext:reg_fe_assrate_robustness} shows average marginal effects of the effect of a home game on the number of physical assaults.\footnote{In nonlinear models, coefficients cannot be interpreted as marginal effects. The partial effect for a Poisson model is given by $\frac{\partial E\left[ \text{Assaults}| \mathbf{X}\right]}{\partial x_j} = \beta_j\ exp(\mathbf{x'}\beta)$. In order to present a single response value, I follow \cite{cameron2005microeconometrics} and report the average response: $\frac{1}{N}\sum_i \frac{\partial E\left[ \text{Assaults}_i| \mathbf{X}_i\right]}{\partial x_{ij}} = \hat\beta_j \times \frac{1}{N} \sum_i\ exp(\mathbf{x}_i'\hat\beta)$.} A home game is predicted to lead to an average increase of 1.483 assaults. This corresponds to an increase in the number of assaults by 28.55 percent. \newline


% Other forms of violence 
\textbf{Other forms of violence.--} In the next step, I investigate the robustness of the findings when looking at other forms of violence. First, a broader definition of assaults is considered. So far, assaults contained offenses coded as `simple bodily harm'. For this robustness test, I further include negligent, dangerous, and grievous bodily harm as well as brawls as assault offenses. Appendix Table \ref{tab_soc_ext:offense_keys_code} illustrates which offenses are contained in the wider assault category and how the penal codes of the German Criminal Code are translated into the offense keys of the Police Crime Statistics. The effect size of 5.417 assaults per million population is significantly larger than the baseline coefficient. This indicates that football games lead to increases in other forms of bodily harm, too. Second, I investigate the effect of football games on threats, which can be regarded as a precursor of physical violence. The effect size with 0.216 threats per million population is relatively small. The estimate corresponds to an increase of 5.4 percent. Third, offenses coded as resistance to enforcement officers are examined. I find that football games lead to, on average, 0.775 offenses per million population. Taken together, the results from this set of robustness checks indicate that other forms of violence are affected by professional football games likewise. This suggests that there are additional violent offenses and no substitution of different crime formats is happening.\newline


% Placebo games
\textbf{Placebo games.--} I estimate the impact of \textit{placebo games} on the assault rate in order to check whether the obtained results are observed just by chance.\footnote{Unfortunately, there are no offenses that may be used as placebo outcomes. This is due to the fact that most of the offenses covered in the PCS are potentially affected by football games.} The true matches take place on roughly 5 percent of the days in the sample. To estimate the effect of placebo games, I drop the affected days with the actual matches and randomly assign dummy indicators with the same frequency of the true matches. Subsequently, I estimate the model as shown in equation \ref{eq_soc_ext:model}. This procedure is carried out 10,000 times and the results are shown in Figure \ref{fig_soc_ext:robustness_placebo_games}. Panel A displays the distribution of the coefficients. As expected, the coefficients are centered around zero. Panel B illustrates the distribution of the t-statistics. The red area under the kernel density indicates significant estimates when using a significance level of $\alpha=0.05$. Panel C shows that 6.57 percent of the estimates are significantly different from zero with 10,000 iterations. With a significance level of $\alpha=0.01$ , there are 3.06 percent significant estimates. The low levels of significant coefficients support the idea that the results obtained earlier are not observed by chance. 



% fig: PLACEBO GAMES
\input{soc_ext_fig__robustness_placebo_games}









%--------------------------------------------------------------------
% CONCLUSION
%-------------------------------------------------------------------
\bigskip
\section{Discussion and Conclusion}\label{sec_soc_ext:conclusion}

% summary
In this paper, I analyze the impact of professional football games on violent behavior. To estimate the causal effects of football games on physical assaults, I use a generalized difference-in-differences approach that exploits the variation in the timing of matches. I compare regional assault levels on days with and without matches conditional on the day of the week, month, and year while additionally accounting for potential confounding variation coming from weather and holidays. I match web-scraped information on 4,461 football matches with data on local assault rates, weather, holidays, and population figures to construct a panel at the municipality-day level over the period 2011-2015. I find that a home game increases the average assault rate by 21.5 percent. The victimization rates of males are driving the results and are particularly high for the age group 18-29 years. Besides, the impact is 
larger for victims with no prior relationship to the suspect and completed offenses. There are no offsetting reductions in assaults on adjacent days to the matchday or in nearby regions. When investigating potential channels, I find no evidence that emotional cues are responsible for the spike of assaults on matchdays. Although there is no support for the frustration-aggression hypothesis, I find large effects for prominent games and teams. In fact, this set of results is suggestive for the fact that spectators mimic players' behavior, select into specific matches, or that the pure agglomeration of fans is causing the higher level of assaults on game days. 




% policy perspective
I find the external effects of football games on violent behavior to be large in magnitude and economically relevant. For instance, professional football games explain 17.7 percent of all assault reports in the regions hosting a professional football club. Back-of-the-envelope calculations indicate that football games in the first three leagues of the German league system precipitate an additional 18,770 assaults in the season of 2014/15,\footnote{These calculations are based on an estimated reduction in assaults per million population per day, 335 days of the football season, and a population of 20,93 million in the affected regions. The number of prevented assaults for the season 2014/15 is: $2.677 \times 335 \times \frac{20,930,000}{1,000,000} = 18,770.$} which imply an annual social cost of 95 million euros.\footnote{To calculate the annual social cost, I use an estimated cost of 5,067 euros (in 2020 prices) for an assault \citep{glaubitz2016kostet}.} The \cite{coase1960problem} theorem states that market failures resulting from externalities can be solved without public intervention (here: the taxpayers do not have to pay for police operations), if there are clearly defined property rights, complete information, and low/no transaction costs. The question is whether the negative externalities would be internalized in absence of state action. To this end, the football clubs (i) must have an incentive to prevent violence in the context of the event and (ii) be able to implement the incentive effectively \citep{daumann2012}. Concerning the former, the clubs have a credible interest in maintaining a peaceful atmosphere otherwise fans, sponsors, and media might avoid the matches. Moreover, the clubs can either employ private security agencies or reimburse police operations to effectively prevent violence in and around the stadiums. Therefore, it should be in the club's own interest to take over the funding of the security. Critics would argue that this procedure is incompatible with the constitution as public safety is a sovereign right and obligation. However, from an economic perspective, it helps to make a distinction between the premises of the football matches and the public domain when discussing the reimbursement of costs for police operations \citep{mause2020}. On the premises, a football match can be considered as a private good, for whose security consumers and producers should pay for.\footnote{The attendance of a football match can be regarded as a private good as both conditions of rivalry (every seat can only be sold once) and excludability (the host can deny entry to some individuals) are satisfied.} In contrast, the socialization of police costs is justifiable in the public domain as all members of the society benefit equally. This regulation, the assumption of police costs on the club's premises, is consistent with statutes in France, Switzerland, and Great Britain. From the point of view of welfare economics, you have to consider both positive and negative externalities and analyze whether football matches increases social welfare. Many positive effects generated by football matches are, however, not externalities in the strict sense as they have already been internalized.\footnote{Many economic agents have to pay a fee in order to be included in the value chain of professional football, such as television companies who pay fees to be allowed to broadcast the sporting events.} Therefore, an interesting point of future research is to assess all relevant externalities, positive and negative alike, and evaluate whether there is a positive social net utility coming from professional football games. 
